---
title: "External Territorial Threats and Tolerance of Corruption: A Private/Government Distinction"
author:
- name: Steven V. Miller
  affiliation: Clemson University
  email: svmille@clemson.edu
appendix: yes
appendixletter: A
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: ~/Dropbox/miscelanea/svm-r-markdown-templates/svm-latex-ms.tex
    toc: yes
    toc_depth: 2
#    citation_package: natbib
#bibliography: "~/Dropbox/master.bib"
fontfamily: mathpazo
fontfamilyoptions: sc, osf
fontsize: 11pt
biblio-style: apsr
subtitle: Supplemental Appendix
tables: yes
anonymous: no
header-includes:
- \usepackage{fontawesome}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, kfigr.prefix=TRUE, kfigr.link=TRUE)
setwd("~/Dropbox/projects/territorial-threats-corruption-attitudes/appendix")
library(ordinal)
library(arm)
library(car)
# library(papeR)
library(tidyverse)
library(countrycode)
library(WDI)
library(QuickUDS)
library(stevemisc)
library(dotwhisker)
library(merTools)
library(stargazer)
```

```{r loaddata, eval=TRUE, echo=FALSE, cache=TRUE, message=F, warning=F}

Data_GB <- read_csv("../analysis/data-gb.csv")
Data_LB <- read_csv("../analysis/data-lb.csv")
Data_WVS <- read_csv("../analysis/data-wvs.csv")

```

\newpage

# Descriptive Statistics

### Descriptive Statistics

Table \ref{tab:summarygb} provides descriptive statistics of the variables used in the Globalbarometer analysis (i.e. Model 1). Table \ref{tab:summarylb} provides descriptive statistics of the variables used in the Latinobarómetro analysis (i.e. Model 2 and Model 3). Table \ref{tab:summarywvs} provides descriptive statistics of the variables used in the World Values Survey analysis. Do note all values are "raw" and that the non-binary independent variables are scaled by two standard deviations in the analyses in the manuscript.

```{r summary_gb, results= "asis", eval=T, echo=F}
Data_GB %>%
  select(corruptmip, age:trust, tarterrmid:initnonterrmid, eti, uds, wip, fotp, ethfrac, sfi, govexp) %>%
  rename(`Corruption is Most Important Problem` = corruptmip,
         `Age` = age,
         `Female` = female,
         `College Educated` = collegeed,
         `Personal Economic Situation` = personalecon,
         `Unemployed` = unemployed,
         `Trust Most People` = trust,
         `Targeted Territorial MID` = tarterrmid,
         `Initiated Territorial MID` = initterrmid,
         `Targeted Non-Territorial MID` = tarnonterrmid,
         `Initiated Non-Territorial MID` = initnonterrmid,
         `Economic Threat Index` = eti,
         `State Fragility Index` = sfi,
         `Level of Democracy` = uds,
         `% of Parliament Held by Women` = wip,
         `Press Freedom` = fotp,
         `Ethnic Fractionalization` = ethfrac,
         `Government Expenditure (% of GDP)` = govexp) %>%
  na.omit() %>% as.data.frame() %>%
  stargazer(., summary=TRUE, type="latex", header=F,
            title = "Summary Statistics for Variables in the Globalbarometer Analysis",
            label="tab:summarygb")
```




```{r summary_lb, results= "asis", eval=T, echo=F}
Data_LB %>%
  select(bribejd, tolgovcord, age:trust, tarterrmid:initnonterrmid, eti, uds, wip, fotp, ethfrac, sfi, govexp) %>%
  rename(`Tolerance of Private Corruption` = bribejd,
         `Tolerance of Government Corruption` = tolgovcord,
         `Age` = age,
         `Female` = female,
         `College Educated` = collegeed,
         `Personal Economic Situation` = socialclass,
         `Unemployed` = unemployed,
         `Trust Most People` = trust,
         `Targeted Territorial MID` = tarterrmid,
         `Initiated Territorial MID` = initterrmid,
         `Targeted Non-Territorial MID` = tarnonterrmid,
         `Initiated Non-Territorial MID` = initnonterrmid,
         `Economic Threat Index` = eti,
         `State Fragility Index` = sfi,
         `Level of Democracy` = uds,
         `% of Parliament Held by Women` = wip,
         `Press Freedom` = fotp,
         `Ethnic Fractionalization` = ethfrac,
         `Government Expenditure (% of GDP)` = govexp) %>%
  as.data.frame() %>%
  stargazer(., summary=TRUE, type="latex", header=F,
            title = "Summary Statistics for Variables in the Latinobarometro Analysis",
            label="tab:summarylb")
```



```{r summary_wvs, results= "asis", eval=T, echo=F}
Data_WVS %>%
  select(bribejd, age:trust, tarterrmid:initnonterrmid, 
         eti, uds, wip, fotp, ethfrac, sfi, govexp) %>%
  rename(`Tolerance of Private Corruption` = bribejd,
         `Age` = age,
         `Female` = female,
         `College Educated` = collegeed,
         `Personal Economic Situation` = satisf,
         `Unemployed` = unemployed,
         `Trust Most People` = trust,
         `Targeted Territorial MID` = tarterrmid,
         `Initiated Territorial MID` = initterrmid,
         `Targeted Non-Territorial MID` = tarnonterrmid,
         `Initiated Non-Territorial MID` = initnonterrmid,
         `Economic Threat Index` = eti,
         `State Fragility Index` = sfi,
         `Level of Democracy` = uds,
         `% of Parliament Held by Women` = wip,
         `Press Freedom` = fotp,
         `Ethnic Fractionalization` = ethfrac,
         `Government Expenditure (% of GDP)` = govexp) %>%
  na.omit() %>%
  as.data.frame() %>%
  stargazer(., summary=TRUE, type="latex", header=F,
            title = "Summary Statistics for Variables in the World Values Survey Analysis",
            label="tab:summarywvs")
```

